Exploring the Feasibility and Performance of\\ One-Step Multi-Factor Authentication with Ear-EEG

### 2018-02-09 paper.tex
Updates from John

\title{Exploring the Feasibility of\\ Multifactor Authentication with EarEEG}

@@

\begin{abstract}
Multifactor authentication presents a robust method to secure our private information, but typically requires multiple actions by the user resulting in a high cost to usability and limiting adoption. A usable system should also be unobtrusive and inconspicuous. We present and discuss a system with the potential to engage all three factors of authentication (inherence, knowledge, and possession) in a single step using an earpiece that implements brain-based authentication using electroencephalography (EEG). We demonstrate its potential by collecting EEG data using manufactured custom-fit earpieces with embedded electrodes and testing a variety of authentication scenarios. Across all participants’ best-performing “passthoughts”, we are able to achieve 0\% false acceptance and 0.36\% false rejection rates, an overall accuracy of 99.82\%, using one earpiece with three electrodes. Furthermore, we find no successful attempts simulating impersonation attacks. We also report perspectives from our participants and discuss directions for future work to explore earEEG for the masses. Our results indicate that a relatively inexpensive system using a single electrode-laden earpiece could provide a discreet, convenient, and robust method for one-step, three-factor authentication.

@@

It is well appreciated by experts and end-users alike that strong authentication is
critical to cybersecurity and privacy, now and into the future. Unfortunately,
news reports of celebrity account hackings serve as regular reminders that
the currently dominant method of authentication in consumer applications, 
single-factor authentication using passwords or other user-chosen secrets, 
faces many challenges. Major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA). However, submitting two different 
authenticators in two separate steps has frustrated wide adoption
due to its additional hassle to users. The Apple iPhone, for instance,
already supports device unlock using either a user-selected passcode or a fingerprint. The
device could very well support a two-step two-factor authentication scheme if
desired. However, it is easy to understand why users would balk at having to
enter a passcode \emph{and} provide a fingerprint each time they want to unlock their phone.

"One-step two-factor authentication" has been proposed as a new approach
to authentication that can provide the security benefits of two-factor authentication without incurring the hassle cost of two-step verification.
In this work we undertake, to the best of our knowledge, the first
ever study and implementation of one-step, \textit{three}-factor authentication. In computer security, authenticators are classified into three types: knowledge factors (e.g., passwords
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics). By taking advantage of a physical token 
in the form of personalized earpieces, the uniqueness of an individual's brainwaves, and
a choice of mental task to use as one's "passthought", we seek to achieve all three factors 
of authentication in a single step by the user.

@@

We find that we can achieve very low false acceptance rates (FAR) and false rejection rates (FRR) with a single, three-electrode earpiece. Interestingly, we found that performance improves in the ear compared to a single electrode on forehead above the left eye, a site typically used for collecting EEG data. We also find that passthoughts could not be spoofed by imposters, even when the imposters had the user's earpiece and knew their chosen passthoughts. These results improve upon previous work on earEEG-based passthoughts systems by testing custom-made earpieces with multiple embedded electrodes which achieve above 99\% performance, an approximately 20\% improvement \cite{curran2016passthoughts}. We also detail important aspects of the system including measurements of the electrodes' electrical impedances and the results of various classifier training and testing schema to probe each authenetication factor. Finally we discuss interesting patterns in our promising results, and future work to investigate the robustness of passthoughts in day-to-day life.

@@

\subsection{Passthoughts and Behavioral Authentication}
The use of EEG as a biometric signal for user authentication has a relatively short history.
In 2005, Thorpe et al. motivated and outlined the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-100\% authentication accuracy for small populations using research-grade and consumer-grade scalp-based EEG systems \cite{Poulos2002,Marcel2007a,Ashby2011,Chuang2013b}.

The concept of in-ear EEG was introduced in 2011 with a demonstration of the feasibility of recording brainwave signals from within the ear canal \cite{Looney2011}. The in-ear placement can produce signal-to-noise ratios comparable to those from conventional EEG electrode placements, is robust to common sources of artifacts, and can be used in a brain-computer interface (BCI) system based on auditory and visual evoked potentials \cite{Kidmose2013a}. One study attempted to demonstrate user authentication using in-ear EEG, but was only able to attain an accuracy level of 80\%, limited by their use of a consumer-grade device with a generic fit, single-electrode earpiece \cite{curran2016passthoughts}. Our study improves on these results by incorporating a physical token (a possession factor) in the form of custom-fitted earpieces with three electrodes each vs. the single channel used in previous work. Furthemore, we use the OpenBCI system which allows us to capture raw data (the device used in previous work used an unknown proprietary algorithm), simultaneously record from the scalp for validation, and report impedance measures important to data quality and authentiation assessment.

@@

When proposing or evaluating authentication paradigms robustness against imposters is often the first consideration, but the usability of these systems is of equal importance as they must conform to a person's needs and lifestyle to warrant adoption and prolonged use. The authors of \cite{sasse2001} describe usability issues with knowledge-based systems like alphanumeric passwords, in particular that it should not be the burden of the user to remember complex passwords that have to be frequently changed. In \cite{braz2006}, researchers analyzed some of the complexities applying interface human factors heuristics to authentication, and indicate the importance of social acceptibility, learnability, and simplicity of authentication methods. Technology worn on the head faces particular usability issues; in their analysis of user perceptions of such devices, the authors of \cite{Genaro2014} identified design, usability, ease of use, and obtrusiveness among the top 10 concerns of users' online postings, and qualitatively comments around comfort and "looking weird".

@@

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability like touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor. \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though it may be vulnerable to "shoulder surfing": imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. Furthermore, this system also presents a possible solution to "rubber-hose attacks" (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the specific performance of which cannot be described (and thus cannot be coerced).

@@

Seven male, right-handed participants (P1-P7), five students and two researchers, were recruited via a university mailing list and completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous and greater diversity is necessary for a larger real-world feasibility assessment beyond this visionary exploration, this quality interestingly functions to strengthen the results of a system designed to discriminate between users (see Discussion). After participants' 3D ear molds were obtained and their fit and electrical impedances were checked, the collection of data for authentication analyses commenced. Data collection consisted of participants completing a demographics questionnaire, a set up period with the OpenBCI system and earpieces with a second impedance check, their performance of nine mental tasks, and finally a post-experiment questionnaire.

@@

We selected a set of mental tasks based on findings in related work regarding the relative strengths of different tasks in authentication accuracy and usability as reported by participants. \cite{Chuang2013b, curran2016passthoughts} Furthermore, given the in-ear placement of the electrodes and therefore the proximity to the temporal lobes containing the auditory cortex, we tested several novel authentication tasks based specifically on aural imagery or stimuli. The nine authentication tasks and their attributes are listed in Table \ref{tab:tasks}. Our strategy was to select tasks that captured a diversity across dimensions of external stimuli, involving a personal secret, eyes open or closed (due to known effects on EEG), and different types of mental imagery.

@@

\subsection{Data Collection Protocol}
%\label{sec:org2041857}
All sites were cleaned with ethanol prior to electrode placement and a small amount of conductive gel was used on each electrode. For EEG recording we used an 8-channel OpenBCI system \cite{michalska2009openbci} which is open-source and costs about 600 USD; an alternative to medical-grade EEG systems (which cost \textgreater20,000 USD), with demonstrated effectiveness \cite{Frey2016}. The ground was placed at the center of the forehead, approximately AFz according to the 10-20 International Standard for Electrode Placement (ISEP), and reference on the left mastoid. The AFz ground location was intentional to not bias left or right ear recordings, though future systems using one ear only should test relocating the ground to a site on one ear (e.g. the earlobe). Six channels were used for the three electrodes on each earpiece (shown in Figure \ref{fig:earpiece_diagram}); for the remaining two channels one AgCl ring electrode was placed on the right mastoid for later re-referencing, and one at approximately Fp1 (ISEP location above the left eye) to validate the data collected in the ears against a common scalp-based placement. Before beginning the experiment, the data from each channel was visually inspected using the OpenBCI interface. Audio stimuli were delivered through small tubes in the earpieces.

@@

We confirm the custom-fit earpieces are able to collect quality EEG data via two metrics: low impedances measured for the ear electrodes, and alpha-band EEG activity attenuation when a participant's eyes were open versus closed.

@@

Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}, to analyze the mental task EEG data. Compared to other linear classifiers, XGBoost uses gradient boosting in which an algorithm generates a decision tree of weak linear classifiers that minimizes a given loss function. Gradient boosting generally improves linear classification results without manually tuning hyperparameters.

To produce feature vectors, we took slices of 100 raw values from each electrode (about 500ms of data), and performed a fourier transform to produce power spectra for each electrode during that slice. We concatenated all electrode power spectra together. No dimensionality reduction was applied. For each task, for each participant, 100 seconds of data were collected in total across 10 trials of 10 seconds each, resulting in 200 samples per participant, per task.

We trained the classifier such that positive examples were from the target participant and target task, and negative examples were randomly selected tasks from any other participant. From this corpus of positive and negative samples, we withheld one third of data for validation. The remaining trainset was used to cross-validate an algorithm over 100 rounds on different splits of the data. The results of each cross-validation (CV) steps were used to iteratively tweak classifier parameters.

For the predictions, the evaluation regards the instances with prediction value larger than 0.5 as positive instances, and the others as negative instances. After updating classifier parameters, the classifier was tested on the withheld validation set. Since negative examples far outweigh positive examples in this dataset, XGBoost atuomatically optimized using the ”error” hyperparameter. Over a set of \(E\) examples containing \(E_W\) wrong examples \(E_W\subset{E}\), XGBoost's binary classification error rate \(\epsilon\) is calculated as

@@

To further test the robustness of the system, we also conducted a ”leave one out” process for the best performing tasks in which each participant’s FAR was calculated once with each other participant left out (e.g. CV for P1 with P2 left out, then CV for P1 with P3 left out, etc. for every participant combination).

@@

For each configuration of electrodes, we calculated the mean FAR and FRR across all participants using each task as the passthought (Figure \ref{fig:meanByElectrode}). Incorporating all electrodes data resulted in the lowest FAR, followed by the combined right and left ear electrodes, respectively. For left ear, right ear, and all electrode configurations every participant had at least one task with 0 FAR/FRR. Counter to our expectations, Fp1 does not perform as well as most ear electrodes, though overall these reported FAR rates are \textless\textless 1\%.

For each position, FAR was about ten times lower than FRR, which is preferable for authentication, as false authentications are more costly than false rejections.

@@

Using only data from the left ear electrodes, the FARs and FRRs of each task for each participant are shown in Tables \ref{tab:farall} and \ref{tab:frrall}, respectively. We find at least one task for each participant that achieves 0\% FAR, and for five participants a task where both the FAR and FRR are 0\%. Each task achieved perfect 0 FAR and FRR for at least one participant, notably \textit{breathe} and \textit{song - open} achieved perfect FAR and FRR for three out of seven participants.

@@

FAR and FRR results by task are shown in Figure \ref{fig:meanByTask}, averaged across participants. Across all tasks, the sport task produced the lowest FAR. Specifically, it produced 0.0\% FAR for all seven participants, with a corresponding 1.8\% FRR. This suggests that the authentication scheme can work very well even if we limit the passthoughts to just a single task category, where the users could choose a personalized secret for that task. Interestingly, tasks like \textit{breathe} and \textit{breathe - open} performed very well despite lacking a personalized secret, indicating that even when the task may be the same across participants our classifier was still able to distinguish between them.

@@

Using our best performing tasks' FARs, averaging 0\% and these tasks' associated FRRs, averaging 0.36\%, we estimate a best-case authentication accuracy of 99.82\%.

@@

These results thus far establish good performance in our default training strategy, in which we count as negative examples recordings from the wrong participant performing any task. For comparison, we report three other analyses with differing negative examples which serve to isolate and test the inherence and knowledge factors: the correct task recorded from the wrong participant (relies on inherence only), the wrong task recorded from the correct participant (relies on knowledge only), and a combination of these two. Positive examples were always the correct participant performing the correct task.

@@

Our "leave one out" analysis with participants' best tasks maintained 0 FAR across all participant combinations.

@@

In addition to the scales and rankings, we included a few open response questions to ascertain attitudes around use cases for in-ear EEG and passthoughts, and the comfort of wearing an in-ear EEG device in everyday life. Participants first read the prompt, "Imagine a commercially available wireless earbud product is now available based on this technology that you've just experienced. It requires minimal effort for you to put on and wear.", and were asked about use cases for in-ear EEG and passthoughts. Responses about in-ear EEG expectedly included authentication for unlocking a phone or computer and building access, but also aspects of self-improvement such as P4's response "Help people increase focus and productivity". P5 and P6 also indicated a use for measuring engagement with media like movies and music, and relatedly P4 wrote "music playback optimized for current mental state and feelings". In terms of comfort wearing such a device, participants generally responded they would be comfortable, though P5 and P6 stipulated only when they already would be wearing something in the ears like earphones. Notably, three participants also added that imaginging a face was difficult and had concerns regarding their ability to repeat tasks in the same exact way each time.

@@

While our authentication analysis establishes that passthoughts achieve low FAR and FRR when tested against other participants' passthoughts, this does not tell us how robust passthoughts are against a spoofing attack, in which both a participant's custom-fit electrode, and details of that participant's chosen passthought, are leaked and authentication is attempted by an imposter.

The first aspect of this scenario we tested was the ability of an imposter to wear an earpiece acquired from someone else and achieve viable impedance values for EEG collection based on the fit of the pieces in their ears. P1 tried on each of the other participants' customized earpieces, which were able to at least physically fit in P1's ears. The impedances of P1 wearing other participants' earpieces were then recorded and are listed in Table \ref{tab:p1_imposter_impedances} below. While some are a better fit than others, overall they are higher (worse) than those achieved by the pieces' intended owners themselves.

@@

To explore the scenario of an imposter attempting to gain access, we chose the case of the most vulnerable participant, P6, whose earpieces P1, P2, and P7 had the lowest impedances while wearing. We collected data using the same data collection protocol, but had the "imposters" refer to P6's report of chosen passsthoughts.
Each imposter performed each of P6's passthoughts (simulating an "inside imposter" from within the system). Following the same analysis steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

Since every participant has one classifier per task (for which that task is the passthought), we are able to make 200 spoofed attempts with the correct passthought on each of P6's classifiers. We find zero successful spoof attempts for tasks with a chosen secret (e.g., \textit{song} or \textit{face}), see Table \ref{tab:imposter_impedances}. In addition, we also do not find any successful spoof attacks for tasks with no chosen secret (e.g., \textit{breathe}). In fact, in all 1,800 spoof attempts (200 attempts for each of the nine classifiers), we do not find a single successful attack on any of P6's classifiers.

However, since this participant's data appeared in the initial pool, the classifier may have been trained on his or her recordings as negative examples. To explore the efficacy of an outsider spoofing recordings, we repeated the same protocol with an individual who did not appear in our initial set of participants (an "outside imposter"). Again, we find zero successful authentications out of 1,800 attempts.

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrr}
& \multicolumn{3}{c}{Impedance [k\(\Omega\)]} \\
\hline
\textbf{P} & \textbf{C} & \textbf{F} & \textbf{B} \\
1 & 18.7 & 10.0 & 8.4\\
2 & 46.7 & 35.7 & 24.8\\
7 & 44.5 & 20.5 & 26.3\\
X & 70.0 & 10.5 & 8.9\\
\end{tabular}
\end{center}
\caption{Left concha (C), canal-front (F) and canal-back (B) electrode impedances of "imposters" P1, P2, P7 and "PX" - a person completely outside of the system - wearing P6's left earpiece.}
\label{tab:imposter_impedances}
\end{table}

>>>>

\title{Exploring the Feasibility and Performance of\\ One-Step 
Multi-Factor Authentication with Ear-EEG}

@@

\begin{abstract}
Multi-factor authentication presents a robust method to secure our private information, but typically requires multiple actions by the user resulting in a high cost to usability and limiting adoption. A usable system should also be unobtrusive and inconspicuous. We present and discuss a system with the potential to engage all three factors of authentication (inherence, knowledge, and possession) in a single step using an earpiece that implements brain-based authentication using electroencephalography (EEG). We demonstrate its potential by collecting EEG data using manufactured custom-fit earpieces with embedded electrodes and testing a variety of authentication scenarios. Across all participants' best-performing ``passthoughts'', we are able to achieve 0\% false acceptance and 0.36\% false rejection rates, for an overall accuracy of 99.82\%, using one earpiece with three electrodes. Furthermore, we find no successful attempts simulating impersonation attacks. We also report on perspectives from our participants. %and discuss directions for future work to explore earEEG for the masses. 
Our results suggest that a relatively inexpensive system using a single electrode-laden earpiece could provide a discreet, convenient, and robust method for one-step multi-factor authentication.

@@

It is well appreciated by experts and end-users alike that strong authentication is
critical to cybersecurity and privacy, now and into the future. Unfortunately,
news reports of celebrity account hackings serve as regular reminders that
the currently dominant method of authentication in consumer applications, 
single-factor authentication using passwords or other user-chosen secrets, 
faces many challenges. Many major online services have strongly encouraged 
their users to adopt two-factor authentication (2FA). However, submitting two 
different authenticators in two separate steps has frustrated wide adoption
due to its additional hassle to users. Modern smartphones, for instance,
already supports device unlock using either a user-selected passcode or a fingerprint. These
devices could very well support a two-step two-factor authentication scheme if
desired. However, it is easy to understand why users would balk at having to
enter a passcode \emph{and} provide a fingerprint each time they want to unlock their phone.

"One-step two-factor authentication" has been proposed as a new approach
to authentication that can provide the security benefits of two-factor authentication 
without incurring the hassle cost of two-step verification \cite{Chuang2014}.
In this work we undertake, to the best of our knowledge, the first-ever study and design of 
one-step, \textit{three}-factor authentication.\footnote{Preliminary findings presented in a non-archival poster at ACM Ubicomp 2017.} 
In computer security, authenticators are 
classified into three types: knowledge factors (e.g., passwords 
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics). By taking advantage of a physical token 
in the form of personalized earpieces, the uniqueness of an individual's brainwaves, and
a choice of mental task to use as one's ``passthought", we seek to achieve all three factors 
of authentication in a single step by the user.

@@

%We find that we can achieve very low false acceptance rates (FAR) and false rejection rates (FRR) with a single, three-electrode earpiece. %Interestingly, we find that performance improves in the ear compared to a single electrode on forehead above the left eye, a site typically used for collecting EEG data. 
%We also find that passthoughts could not be spoofed by imposters, even when the imposters had the user's earpiece and knew their chosen passthoughts. These results improve upon previous work on ear-EEG based passthoughts systems by testing custom-made earpieces with multiple embedded electrodes which achieve above 99\% performance, an approximately 20\% improvement \cite{curran2016passthoughts}. We also detail important aspects of the system including measurements of the electrodes' electrical impedances and the results of various classifier training and testing schema to probe each authentication factor. Finally we discuss interesting patterns in our promising results, and future work to investigate the robustness of passthoughts in day-to-day life.
%
%
%
%
We make the following contributions in this work. First, we achieve a 99.82\% authentication accuracy with zero False Acceptance Rate (FAR) using personalized custom-fit three-channel EEG earpieces and a passthoughts authentication paradigm. Second, we quantify the improvements over prior art in authentication accuracy due to the use of custom-fit versus generic earpieces, and the use of multiple electrodes versus a single electrode. Third, we evaluate multiple classification strategies that allows us to compare the relative contributions of the inherence factor and knowledge factor to authentication accuracy. Fourth, we perform simulation attacks to demonstrate the method's robustness against impersonation via four scenarios where the attacker have access to the target's earpiece and/or secret passthoughts. Collectively, we build a case that passthoughts authentication using personalized custom-fit earpieces offers a viable and attractive path towards one-step three-factor authentication.

@@

\subsection{Passthoughts and Behavioral Authentication}
The use of EEG as a biometric signal for user authentication has a relatively short history.
In 2005, Thorpe et al. motivated and outlined the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-100\% authentication accuracy for small populations using research-grade and consumer-grade scalp-based EEG systems \cite{Poulos2002,Marcel2007a,Ashby2011,Chuang2013b}. Several recent works on brainwave biometrics have independently demonstrated individuals' EEG permanence over one to six months \cite{Armstrong2015,Maiorana2016} or even over one year \cite{Ruiz2017}.

The concept of in-ear EEG was introduced in 2011 with a demonstration of the feasibility of recording brainwave signals from within the ear canal \cite{Looney2011}. The in-ear placement can produce signal-to-noise ratios comparable to those from conventional EEG electrode placements, is robust to common sources of artifacts, and can be used in a brain-computer interface (BCI) system based on auditory and visual evoked potentials \cite{Kidmose2013a}. 
%
In a previous study, we 
%A previous study 
attempted to demonstrate user authentication using in-ear EEG, but were only able to attain an accuracy level of 80\%, limited by the use of a consumer-grade device with a single generic-fit electrode \cite{curran2016passthoughts}.
%In this present study, we show that by incorporating a physical token in the form of a custom-fitted earpiece with three electrodes, we can achieve a significant improvement of accuracy to 99.82\%.

%Our study significantly improves on these results by incorporating a physical token (a possession factor) in the form of custom-fitted earpieces with three electrodes each vs. the single channel used in previous work. 
%Furthermore, we use the OpenBCI system which allows us to capture raw data (the device used in previous work used an unknown proprietary algorithm), 
%Furthermore, we simultaneously record from the scalp for validation, and report impedance measures important to data quality and authentication assessment.

@@

When proposing or evaluating authentication paradigms, robustness against imposters is often a first consideration, but the usability of these systems is of equal importance as they must conform to a person's needs and lifestyle to warrant adoption and prolonged use. The authors of \cite{sasse2001} describe usability issues with knowledge-based systems like alphanumeric passwords, in particular that it should not be the burden of the user to remember complex passwords that have to be frequently changed. In \cite{braz2006}, researchers analyzed some of the complexities applying interface human factors heuristics to authentication, and indicate the importance of social acceptability, learnability, and simplicity of authentication methods. Technologies worn on the head face particular usability issues; in their analysis of user perceptions of such devices, the authors of \cite{Genaro2014} identified design, usability, ease of use, and obtrusiveness among the top ten concerns of users, as well as qualitative comments around comfort and ``looking weird''.

@@

Mobile and wearable technologies' continuous proximity to the user's body provides favorable conditions for unobtrusively capturing biometrics for authentication. Many such uses have been proposed that embrace usability like touch-based interactions \cite{Tartz2015, Holz2015} and walking patterns \cite{Lu2014} using mobile phones, as well as identification via head movements and blinking in head-worn devices \cite{Rogers2015}. However, these typically draw only from the inherence factor. \cite{Chen2015} proposed an inherence and knowledge two-factor method for multi-touch mobile devices based on a user's unique finger tapping of a song, though it may be vulnerable to ``shoulder surfing'': imposters observing and mimicking the behavior. By comparison, passthoughts are largely invisible to attackers. Furthermore, this system also presents a possible solution to ``rubber-hose attacks'' (forceful coercion to divulge a password): a thought, which is secret (and thus changeable), but has a particular expression unique to an individual, the specific performance of which cannot be described (and thus cannot be coerced).

@@

Seven male, right-handed participants (P1-P7), five students and two researchers, were recruited via a university mailing list and completed our study protocol approved by our local ethics review board. Though this sample is relatively homogenous and greater diversity is necessary for a larger real-world feasibility assessment% beyond this visionary exploration
, this quality interestingly functions to strengthen the results of a system designed to discriminate between users (see Discussion). After participants' 3D ear molds were obtained, the custom-fit earpieces were manufactured, and their fit and electrical impedances were checked, we proceeded to the collection of study data. %for authentication analyses commenced. 
Data collection consisted of participants completing a demographics questionnaire, a setup period with the OpenBCI system and earpieces with a second impedance check, their performance of nine mental tasks, and finally a post-experiment questionnaire.

@@

We selected a set of mental tasks based on findings in related work regarding the relative strengths of different tasks in authentication accuracy and usability as reported by participants \cite{Chuang2013b, curran2016passthoughts}. Furthermore, given the in-ear placement of the electrodes and therefore the proximity to the temporal lobes containing the auditory cortex, we tested several novel authentication tasks based specifically on aural imagery or stimuli. The nine authentication tasks and their attributes are listed in Table \ref{tab:tasks}. Our strategy was to select tasks that captured a diversity across dimensions of external stimuli, involving a personal secret, eyes open or closed (due to known effects on EEG), and different types of mental imagery.

@@

\subsection{Data Collection Protocol}
%\label{sec:org2041857}
All sites were cleaned with ethanol prior to electrode placement and a small amount of conductive gel was used on each electrode. For EEG recording we used an 8-channel OpenBCI system \cite{michalska2009openbci} which is open-source and costs about 600 USD; an alternative to medical-grade EEG systems (which cost \textgreater20,000 USD), with demonstrated effectiveness \cite{Frey2016}. The ground was placed at the center of the forehead, at AFz according to the 10-20 International Standard for Electrode Placement (ISEP), and reference on the left mastoid. The AFz ground location was intentional to not bias left or right ear recordings, though future systems using one ear only should test relocating the ground to a site on one ear (e.g., the earlobe). Six channels were used for the three electrodes on each earpiece (shown in Figure \ref{fig:earpiece_diagram}). For the remaining two channels, one AgCl ring electrode was placed on the right mastoid for later re-referencing, and one at Fp1 (ISEP location above the left eye) to validate the data collected in the ears against a common scalp-based placement. Before beginning the experiment, the data from each channel was visually inspected using the OpenBCI interface. Audio stimuli were delivered through small tubes in the earpieces.

@@

We confirm that the custom-fit earpieces are able to collect quality EEG data via two metrics: low impedances measured for the ear electrodes, and alpha-band EEG activity attenuation when a participant's eyes were open versus closed.

@@

Since past work has shown that classification tasks in EEG-based brain-computer interfaces (BCI) are linear \cite{Garrett2003a}, we used XGBoost, a popular tool for logistic linear classification \cite{Chen2016}, to analyze the mental task EEG data. Compared to other linear classifiers, XGBoost uses gradient boosting in which an algorithm generates a decision tree of weak linear classifiers that minimizes a given loss function. Gradient boosting generally improves linear classification results without manually tuning hyper-parameters.

To produce feature vectors, we took slices of 100 raw values from each electrode (about 500ms of data), and performed a Fourier transform to produce power spectra for each electrode during that slice. We concatenated all electrode power spectra together. No dimensionality reduction was applied. For each task, for each participant, 100 seconds of data were collected in total across 10 trials of 10 seconds each, resulting in 200 samples per participant, per task.

We trained the classifier such that positive examples were from the target participant and target task, and negative examples were randomly selected tasks from any other participant. From this corpus of positive and negative samples, we withheld one third of data for validation. The remaining training set was used to cross-validate an algorithm over 100 rounds on different splits of the data. The results of each cross-validation (CV) steps were used to iteratively tweak classifier parameters.

For the predictions, the evaluation regards the instances with prediction value larger than 0.5 as positive instances, and the others as negative instances. After updating classifier parameters, the classifier was tested on the withheld validation set. Since negative examples far outweigh positive examples in this dataset, XGBoost automatically optimized using the €error€ hyperparameter. Over a set of \(E\) examples containing \(E_W\) wrong examples \(E_W\subset{E}\), XGBoost's binary classification error rate \(\epsilon\) is calculated as

@@

To further test the robustness of the system, we also conducted a `€leave one out' process for the best performing tasks in which each participant's FAR was calculated once with each other participant left out (e.g., CV for P1 with P2 left out, then CV for P1 with P3 left out, etc., for every participant combination).

@@

For each configuration of electrodes, we calculated the mean FAR and FRR across all participants using each task as the passthought (Figure \ref{fig:meanByElectrode}). Incorporating all electrodes data resulted in the lowest FAR, followed by the combined right and left ear electrodes, respectively. For left ear (3 electrodes), right ear (3 electrodes), and both ears (6 electrodes) configurations, every participant had at least one task with zero FAR and FRR. Among the individual electrodes, the left canal front electrode produced a mean FAR of 0.12\% and a mean FRR just below 20\%. Counter to our expectations, Fp1 does not perform as well as most ear electrodes, though overall these reported FAR rates are \textless\textless 1\%.

For each position, FAR was about ten times lower than FRR, which is preferable for authentication, as false authentications are generally more costly than false rejections.

@@

Using only data from the three left ear electrodes, the FARs and FRRs of each task for each participant are shown in Tables \ref{tab:farall} and \ref{tab:frrall}, respectively. We find at least one task for each participant that achieves 0\% FAR, and for five participants a task where both the FAR and FRR are 0\%. Each task achieved perfect 0\% FAR and FRR for at least one participant, notably \textit{breathe} and \textit{song - open} achieved perfect FAR and FRR for three out of seven participants.

@@

FAR and FRR results by task are shown in Figure \ref{fig:meanByTask}, averaged across participants. Across all tasks, the sport task produced the lowest FAR. Specifically, it produced 0\% FAR for all seven participants, with a corresponding 1.8\% FRR. This suggests that the authentication scheme can work very well even if we limit the passthoughts to just a single task category, where the users could choose a personalized secret for that task. Interestingly, tasks like \textit{breathe} and \textit{breathe - open} performed very well despite lacking a personalized secret, indicating that even when the task may be the same across participants our classifier was still able to distinguish between them.

@@

Using our best performing tasks' FARs, averaging 0\% and these tasks' associated FRRs, averaging 0.36\%, we obtain an authentication accuracy of 99.82\% when we use the three electrodes in the left ear. For comparison, if we limit ourselves to a single electrode (left canal front), we obtain an authentication accuracy of 90\%.

@@

Our ``leave one out'' analysis with participants' best tasks maintained 0\% FAR across all participant combinations.

\subsubsection{Relative Contributions of Inherence and Knowledge Factors}

These results thus far establish good performance in our default training strategy, in which we count as negative examples recordings from the wrong participant performing any task. We further performed three other analyses with differing negative examples which serve to isolate and test the inherence and knowledge factors: the correct task recorded from the wrong participant (relies on inherence only), the wrong task recorded from the correct participant (relies on knowledge only), and a combination of these two. Positive examples were always the correct participant performing the correct task.

@@

@@

In addition to the scales and rankings, we included a few open response questions to ascertain attitudes around use cases for in-ear EEG and passthoughts, and the comfort of wearing an in-ear EEG device in everyday life. Participants first read the prompt, "Imagine a commercially available wireless earbud product is now available based on this technology that you've just experienced. It requires minimal effort for you to put on and wear.", and were asked about use cases for in-ear EEG and passthoughts. Responses about in-ear EEG expectedly included authentication for unlocking a phone or computer and building access, but also aspects of self-improvement such as P4's response "Help people increase focus and productivity". P5 and P6 also indicated a use for measuring engagement with media like movies and music, and relatedly P4 wrote "music playback optimized for current mental state and feelings". In terms of comfort wearing such a device, participants generally responded they would be comfortable, though P5 and P6 stipulated only when they already would be wearing something in the ears like earphones. Notably, three participants also added that imagining a face was difficult and had concerns regarding their ability to repeat tasks in the same exact way each time.

@@

While our authentication analysis establishes that passthoughts achieve low FAR and FRR when tested against other participants' passthoughts, this does not tell us how robust passthoughts are against a spoofing attack, in which both a participant's custom-fit earpiece, and details of that participant's chosen passthought, are leaked and authentication is attempted by an imposter. We perform four different analyses to investigate the system's robustness against imposter attacks.

First, we test the ability of an imposter to wear an earpiece acquired from someone else and achieve viable impedance values for EEG collection based on the fit of the pieces in their ears. P1 tried on each of the other participants' customized earpieces.%, which were able to at least physically fit in P1's ears. 
The impedances from each electrode were recorded and are listed in Table \ref{tab:p1_imposter_impedances} below. 
%The impedances of P1 wearing other participants' earpieces were then recorded and are listed in Table \ref{tab:p1_imposter_impedances} below. 
Across all cases, the impedances are not only higher (worse), but also deviate significantly from those achieved by the pieces' intended owners themselves (Table \ref{tab:impedances}).
%While some are a better fit than others, overall they are significantly higher (worse) than those achieved by the pieces' intended owners themselves (Table \ref{tab:impedances}).
These results come as no surprise given the uniqueness of ear canal shapes between individuals \cite{Akkermans2005}, and point to the possibility that the presentation of a physical token that provides the correct impedance levels can be used as another demonstration of both the inherence and possession factors.
%While these results are based on only one participant, they nonetheless point to the possibility that the presentation of a physical token that provides the correct impedance levels can be used as another demonstration of both the inherence and possession factors.

@@

Second, to explore the scenario of an imposter attempting to gain access, we choose the case of the most vulnerable participant, P6, whose earpieces P1, P2, and P7 had the lowest impedances while wearing (Table \ref{tab:imposter_impedances}). We collected data using the same data collection protocol, but had the "imposters" refer to P6's report of chosen passsthoughts.
Each imposter performed each of P6's passthoughts (simulating an "inside imposter" from within the system). Following the same analysis steps, we generated 200 samples per task for our imposters, using data from all left ear electrodes.

@@

Since every participant has one classifier per task (for which that task is the passthought), we are able to make 200 spoofed attempts with the correct passthought on each of P6's classifiers. We find zero successful spoof attempts for tasks with a chosen secret (e.g., \textit{song} or \textit{face}). In addition, we also do not find any successful spoof attacks for tasks with no chosen secret (e.g., \textit{breathe}). In fact, in all 1,800 spoof attempts (200 attempts for each of the nine classifiers), we do not find a single successful attack on any of P6's classifiers.

Since this participant's data appeared in the initial pool, the classifier may have been trained on his or her recordings as negative examples. As our third analysis, to explore the efficacy of an outsider spoofing recordings, we repeated the same protocol with an individual ``PX'' who did not appear in our initial set of participants (an ``outside imposter''). Again, we find zero successful authentications out of 1,800 attempts.

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrr}
& \multicolumn{3}{c}{Impedance [k\(\Omega\)]} \\
\hline
\textbf{P} & \textbf{C} & \textbf{F} & \textbf{B} \\
1 & 18.7 & 10.0 & 8.4\\
2 & 46.7 & 35.7 & 24.8\\
7 & 44.5 & 20.5 & 26.3\\
X & 70.0 & 10.5 & 8.9\\
\end{tabular}
\end{center}
\caption{Left concha (C), canal-front (F) and canal-back (B) electrode impedances of ``imposters'' P1, P2, P7 and ``PX'' - a person completely outside of the system - wearing P6's left earpiece.}
\label{tab:imposter_impedances}
\end{table}

Fourth, our ``leave one out'' analysis can also be seen as another set of outside imposter attacks, in which each participant acts as an outside imposter for each other participant, but where the imposters have their own manufactured earpieces and passthoughts. The best task classifiers achieved FARs of 0\% across all combinations, successfully rejecting the simulated imposters.

\section{Discussion, Limitations, \& Future Work}
